Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-21 paper_jasa.tex
figure related references updated

\hh{From figure \ref{fig:power_loess3} we see that identification of the data plot is high independently of the effect size. This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e. the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.} \green{Not sure that this goes here, but something needs to be said about Type III.}

@@

In the real setting, where visual inference is to be useful, there will be no conventional tests or traditional $p$-values. Yet, assessing the strength of the structure is a critical component of inference. For experiments 1 and 2, to assess the signal strength we examine the proportion of correct picks, with multiple users and the $p$-value of the actual data plot. For experiment 3 the \green{How is p-value calculated here?} Figure \ref{fig:pval_pcorrect} displays scatterplots of these two variables for each experiment. As the $p$-value increases the proportion of correct responses falls, evidence of direct association between proportion of correct responses and conventional test $p$-values. When $p$-value is larger than 0.15, it is almost impossible to correctly identify the observed plot in the lineup. We observed this for both experiment 1 and 2. For experiment 3, the scenario is different, because there are two structures, the association between variables, and the cluster of contaminated points.% as in this case people might have observed something other than the effect $E$.

@@

In Section \ref{sec:size} in order to make comparisons with conventional tests an assumption was made that the observer would pick the plot in the lineup that had the lowest $p$-value. In this section we take a look at the picks that subjects made, to see if this is, indeed, true. Figure \ref{fig:P-val_rank} shows side-by-side boxplots of rank of plot (in its lineup) and how many subjects selected the plot. \green{Maybe we don't need this plot.}

@@

See figure \ref{still-to-do} of evidence for this phenomenon.
Mahbub, could you have a look at the lineups, and compute the difference between the data plot's $p$-value and the  smallest $p$ value of a null plot, and plot that in a scatterplot against proportion correct? Thanks!

This problem is further investigated in the paper \cite{waldo}. \blue{Here is the plot in figure \ref{fig:pval_difference} as proposed.}

>>>>

\hh{From figure \ref{fig:power_loess_effect} we see that identification of the data plot is high independently of the effect size for experiment 3. This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e. the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.} \green{Not sure that this goes here, but something needs to be said about Type III.}

@@

In the real setting, where visual inference is to be useful, there will be no conventional tests or traditional $p$-values. Yet, assessing the strength of the structure is a critical component of inference. For experiments 1 and 2, to assess the signal strength we examine the proportion of correct picks, with multiple users and the $p$-value of the actual data plot. For experiment 3 the \green{How is p-value calculated here?} Figure \ref{fig:pval_pcorrect} \blue{(should I bring this figure back?)}  displays scatterplots of these two variables for each experiment. As the $p$-value increases the proportion of correct responses falls, evidence of direct association between proportion of correct responses and conventional test $p$-values. When $p$-value is larger than 0.15, it is almost impossible to correctly identify the observed plot in the lineup. We observed this for both experiment 1 and 2. For experiment 3, the scenario is different, because there are two structures, the association between variables, and the cluster of contaminated points.% as in this case people might have observed something other than the effect $E$.

@@

In Section \ref{sec:size} in order to make comparisons with conventional tests an assumption was made that the observer would pick the plot in the lineup that had the lowest $p$-value. In this section we take a look at the picks that subjects made, to see if this is, indeed, true.

@@

See figure \ref{fig:pval_difference} of evidence for this phenomenon.
%Mahbub, could you have a look at the lineups, and compute the difference between the data plot's $p$-value and the  smallest $p$ value of a null plot, and plot that in a scatterplot against proportion correct? Thanks!

This problem is further investigated in the paper \cite{waldo}.
